---
title: "Coursera Machine Learning Project"
author: "Kevin Graebel"
date: "December 3, 2018"
output: html_document
---
### Loading of full data set

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, it is necessary to load the packages needed.
```{r, echo=TRUE, message = FALSE, warning = FALSE}
library(dplyr)
library(caret)
library(ggplot2)
library(data.table)
library(randomForest)
```


```{r, warning = FALSE, cache=TRUE, cache.lazy=FALSE}
pmlTrain <- read.csv(file = "pml-training.csv", header = TRUE)
pmlTest <- read.csv(file = "pml-testing.csv", header = TRUE)
```

First, We will eliminate columns that are mostly NA.

```{r}
naCols <- which(apply(pmlTrain,2,anyNA))
pmlTrain <- pmlTrain[, -naCols]
pmlTest <- pmlTest[, -naCols]
```

Next, we can strip out the variables with near zero variability to eliminate
unnecessary covariates, including the timestamp data at the front.

```{r}
nzv <- nearZeroVar(pmlTrain)
pmlTrain <- pmlTrain[,-nzv]
pmlTest <- pmlTest[,-nzv]
pmlTrain <- pmlTrain[,-c(1:6)]
pmlTest <- pmlTest[,-c(1:6)]
```

Now, we will split into a training and validation set to test the accuracy of the model 
before applying it to the actual testing set of the quiz.
```{r}
intrain <- createDataPartition(pmlTrain$classe, p = 0.75, list = FALSE)
pmlTrain1 <- pmlTrain[intrain,]
pmlValid1 <- pmlTrain[-intrain,]
```

Now, we can build a random forest model and review the accuracy of the confusion matrix
compared to the validation training set.

```{r}
modelRF <- randomForest(classe ~., data = pmlTrain1)
predRF <- predict(modelRF, pmlValid1)
confusionMatrix(pmlValid1$classe, predRF)
#table(predict(modFit, pmlTest), pmlTest$classe)
```

Given that the results of this look strong, we can apply this model to the testing 
data set to answer the quiz questions
```{r}
predict(modelRF, pmlTest)
```


###Citation of the dataset and literature review.
Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 